---
title: "Covid91 vaccine study Final2022"
author: "Kaylie Brehm"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
Male = filter(FinalData, sex =="M")
Female = filter(FinalData, sex =="F")
Gay = filter(FinalData, LGBTQ =="gay")
DrugUser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.



# Hypothesis and Method

$H_0$: The preventative vaccine, covid91, is not effective for this particular subgroup (male, female, gay, or drug user).

$H_A$: The preventative vaccine, covid91, is effective for this particular subgroup (male, female, gay, or drug user).

Whether an individual was infected or not is categorical. Whether an individual was male or female is categorical. Whether an individual is gay or straight is categorical. Whether an individual is a drug user or not is categorical. I will use a CAT~CAT analysis to determine whether the vaccine was effective for each particular subgroup. In each subgroup section, I will use a bar chart (graphical descriptive results), a data table (numerical descriptive results), a chi-squared test (inferential descriptive results), and a fisher exact test (inferential descriptive results). I will also define the upper and lower confidence intervals, and run a test for effectiveness. Then, in each subgroup section, I will analyze the data and give a conclusion. Once each subgroup is completed, I will give my overall thoughts thoughts on the study. 

## Male

### Graphical Descriptive Results

```{r}
dd2 <- Male %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The plots represent the population of males and their relationship to getting covid-19 by taking either the vaccine or placebo. There are two plots that came out. Both plots show two bars for treatment, the placebo or the vaccine. Along the x-axis is the count of people who came in contact with covid. Along the y-axis is those infected with covid. The first plot shows the number of people infected with covid. The second plot shows the percentage of people infected with covid.



### Numerical Descriptive Results 

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

### Inferential Descriptive Results 

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
chisq.test(table1)
chisqtestGC(table1)
malesfish <- fisher.test(table1)
malesfish
```

### Effictiveness and Confidence Intervals

```{r}
effect <- malesfish$p.value/(malesfish$p.value +1)
effectivenessguy = malesfish$estimate/(malesfish$estimate +1)*100
print(effectivenessguy)
```


```{r}
low=(malesfish$conf.int[1]/(malesfish$conf.int[1] +1))*100
hi=(malesfish$conf.int[2]/(malesfish$conf.int[2] +1))*100
options(digits=2)
```

### Conclusion



## Female

### Graphical Descriptive Results

```{r}
dd2 <- Female %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The plots represent the population of females and their relationship to getting covid-19 by taking either the vaccine or placebo. There are two plots that came out. Both plots show two bars for treatment, the placebo or the vaccine. Along the x-axis is the count of people who came in contact with covid. Along the y-axis is those infected with covid. The first plot shows the number of people infected with covid. The second plot shows the percentage of people infected with covid.

### Numerical Descriptive Results 

```{r}
table2 <- xtabs(~infected + treatment, data=Female)
rowPerc(table2)
colPerc(table2)
```

### Inferential Descriptive Results 

```{r}
table2 <- xtabs(~infected + treatment, data=Female)
chisq.test(table2)
chisqtestGC(table2)
femalesfish <- fisher.test(table2)
femalesfish
```

### Effictiveness and Confidence Intervals

```{r}
effect <- femalesfish$p.value/(femalesfish$p.value +1)
effectivenessfemales = femalesfish$estimate/(femalesfish$estimate +1)*100
print(effectivenessfemales)
```

```{r}
low=(femalesfish$conf.int[1]/(femalesfish$conf.int[1] +1))*100
hi=(femalesfish$conf.int[2]/(femalesfish$conf.int[2] +1))*100
options(digits=2)
```

### Conclusion



## Gay

### Graphical Descriptive Results

```{r}
dd2 <- Gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The plots represent the population of gays and their relationship to getting covid-19 by taking either the vaccine or placebo. There are two plots that came out. Both plots show two bars for treatment, the placebo or the vaccine. Along the x-axis is the count of people who came in contact with covid. Along the y-axis is those infected with covid. The first plot shows the number of people infected with covid. The second plot shows the percentage of people infected with covid.

### Numerical Descriptive Results 

```{r}
table3 <- xtabs(~infected + treatment, data=Gay)
rowPerc(table3)
colPerc(table3)
```

### Inferential Descriptive Results 

```{r}
table3 <- xtabs(~infected + treatment, data=Gay)
chisq.test(table3)
chisqtestGC(table3)
gayfish <- fisher.test(table3)
gayfish
```



### Effictiveness and Confidence Intervals

```{r}
effect <- gayfish$p.value/(gayfish$p.value +1)
effectivenessgay = gayfish$estimate/(gayfish$estimate +1)*100
print(effectivenessgay)
```

```{r}
low=(gayfish$conf.int[1]/(gayfish$conf.int[1] +1))*100
hi=(gayfish$conf.int[2]/(gayfish$conf.int[2] +1))*100
options(digits=2)
```


### Conclusion



## Drug User

### Graphical Descriptive Results

```{r}
dd2 <- DrugUser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The plots represent the population of drug users and their relationship to getting covid-19 by taking either the vaccine or placebo. There are two plots that came out. Both plots show two bars for treatment, the placebo or the vaccine. Along the x-axis is the count of people who came in contact with covid. Along the y-axis is those infected with covid. The first plot shows the number of people infected with covid. The second plot shows the percentage of people infected with covid.

### Numerical Descriptive Results 

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table4)
colPerc(table4)
```

### Inferential Descriptive Results 

```{r}
table4 <- xtabs(~infected + treatment, data=DrugUser)
chisq.test(table4)
chisqtestGC(table4)
Druguserfish <- fisher.test(table4)
Druguserfish
```

### Effictiveness and Confidence Intervals

```{r}
effect <- Druguserfish$p.value/(Druguserfish$p.value +1)
effectivenessDruguser = Druguserfish$estimate/(Druguserfish$estimate +1)*100
print(effectivenessDruguser)
```

```{r}
low=(Druguserfish$conf.int[1]/(Druguserfish$conf.int[1] +1))*100
hi=(Druguserfish$conf.int[2]/(Druguserfish$conf.int[2] +1))*100
options(digits=2)

```


### Conclusion



# Final Thoughts and Discussion

